---
title: "Statistical Inference - final project"
author: "Wiktoria Urantowka"
date: "8/11/2017"
output:
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
The goal of this project is the applications of Central Limit THeorem to  exponential distribution in R. The last one is simulated in R using command rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is the same as it's standard deviation and amounts 1/lambda.

##A. Simulations##
1. setting constants
```{r}
lambda <- 0.2 # lambda
n <- 40 # size of the sample
sim_num <- 1000 # number of drows
```
2. Setting seed for reproducibility
```{r}
set.seed(67893)
```   
3. Simulating values from exponential distribution
```{r}
sim_values<- rexp(n = sim_num * n,lambda)
hist(sim_values, main = "simulated values", breaks =200)
``` 

4. Create a dataframe with means of 1000 simulation of size n=40
```{r}
matrix <- matrix(rexp(n = sim_num * n,lambda), sim_num, n)
means <- rowMeans(matrix)
simData <- data.frame(cbind(matrix,means))
```


##B. Comparaison of theoretical and sample means of the distribution.##

a. Theoretical mean of exponential distribution = 1/lambda
```{r}
theoretical_mean <- 1/lambda
theoretical_mean
```
b. Mean from one drow of n=40 is given by:
```{r}
sample_mean <-mean(rexp(40, 0.2))
sample_mean
```
and the difference between the two is small
```{r}
mean_difference<-abs(theoretical_mean-sample_mean)
mean_difference
```

c. However, when we repeat it 1000 times and compute mean,
```{r}
mns = NULL
for (i in 1 : 1000) mns = c(mns, mean(rexp(40, 0.2)))
sample_mean1000<- mean(mns)
mean_difference2<-theoretical_mean-sample_mean1000
sample_mean1000
mean_difference2
```
One can see, that the difference between sample mean and theoretical mean shrinks further  

d. Plotting the results and finding their mean value
```{r}
library("ggplot2")
ggplot(data = simData, aes(simData$means)) + 
  geom_histogram(breaks = seq(2, 9, by = 0.2)) + 
  labs(title = "Histogram of the Means", x = "Means", y = "Frequency") + 
  geom_vline(aes(xintercept=mean(simData$means)), color="red", 
 size=2)
```
The histogram proves that the distribution of the means is centered around the mean of the simulated distribution, very very close to the theoretical mean = 5

##2. Comparaison of theoretical and sample variance of the distribution.##

a. Theoretical variance of exponential distribution is given by: 
```{r}
theoretical_variance = ((1 / lambda) ^ 2) / n
theoretical_variance
```
b. sample variance is given by
```{r}
sample_variance<-var(means)
sample_variance

var_difference <- theoretical_variance-sample_variance
var_difference
```
One can see, that the variance difference between the theoretical and sample variance is very small   

##Show that the distribution is approximately normal.##
Showing that the simulated mean sample data approximately follows the Normal distribution when n is big (Central Limit Theorem) can be illustrated below:
(Comparison of the plotted sample data with the plot of normal distribution) 
```{r}
qplot(means, geom = 'blank') + 
  geom_line(aes(y=..density.., colour='Empirical'), stat='density', size=1) + 
  stat_function(fun=dnorm, args=list(mean=(1/lambda), sd=((1/lambda)/sqrt(n))), 
                aes(colour='Normal'), size=1) + 
  geom_histogram(aes(y=..density.., fill=..density..), alpha=0.4, 
                 breaks = seq(2, 9, by = 0.2), col='red') + 
  scale_fill_gradient("Density", low = "yellow", high = "red") + 
  scale_color_manual(name='Density', values=c('brown', 'blue')) 
```   

  Comparying the "Empirical" (derived from exponential distribution) and "Normal" distributions clearly illustrates, that the sample data can be approximated with the normal distribution. 
  
  
  