---
title: "Tooth Growth Analysis"
author: "Wiktoria Urantowka"
date: "8/20/2017"
output:
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The goal of this report id to analyze the ToothGrowth data in R in order to investigate the impact of vitamine C on the tooth growth in ginea pigs. 

###1.Loading the data and basic exploratory analysis###
```{r}
data(ToothGrowth)
head(ToothGrowth)
dim(ToothGrowth)
str(ToothGrowth)
```
The data consists of 60 observations for 3 variables: Tooth length (numeric), supplement type, VC or OJ (factor) and the dosage in milligrams of this supplement (numeric) 
```{r}
library(ggplot2)
qplot(supp,len,data=ToothGrowth, facets=~dose, main="Tooth growth by supplement type and dosage (mg)",xlab="Supplement", ylab="Tooth length") + geom_boxplot(aes(fill = supp))
```
 
One can observe a positive correlation between the dosage and the tooth growth.
For lower dosage, the affect of VC seams weaker than OJ, but this difference dissapears with the highest dosage. 

###2.Hypothesis Testing:###
#### 1 Difference of effect between supplements####
Ho =  no difference in tooth growth when using the supplement OJ and VC (lenOJ == lenVC)
Ha = Tooth grow more when using supplement OJ instead of VC. (lenOJ>lenVC)

Assumptions
-Variances of tooth growth are different for different supplements and their dosage.
-Tooth growth follows normal distribution.
-variables are i.i.d.

Procedure:
- tooth growth by supplement type from the data
```{r}
OJ = ToothGrowth$len[ToothGrowth$supp == 'OJ']
VC = ToothGrowth$len[ToothGrowth$supp == 'VC']
```

-One-tailed independent t-test with unequal variance.
```{r}
t.test(OJ, VC, alternative = "greater", paired = FALSE, var.equal = FALSE, conf.level = 0.95)
```
As the p-value < 0.05 the null hypothesis is rejected. 
--> Supplement OJ has bigger effect on tooth growth than supplement VC

#### 2 Difference of effects among dosages####
2.1(Comparaison between affect of dosage 0.5 and 1)
Ho=  no difference in tooth growth between dosages. 
Ha= more tooth growth when the dosage increases.

Procedure:
-tooth growth by dosage.
```{r}
doseHalf = ToothGrowth$len[ToothGrowth$dose == 0.5]
doseOne = ToothGrowth$len[ToothGrowth$dose == 1]
doseTwo = ToothGrowth$len[ToothGrowth$dose == 2]
```

-One-tailed independent t-test with unequal variance.
```{r}
t.test(doseHalf, doseOne, alternative = "less", paired = FALSE, var.equal = FALSE, conf.level = 0.95)
```
As the p-value < 0.05 the null hypothesis is rejected. (There is a difference in effects on tooth growth for dosages 0.5 and 1, dosage 1 having bigger effect)

2.2(Comparaison between affect of dosage 1 and 2)
```{r}
t.test(doseOne, doseTwo, alternative = "less", paired = FALSE, var.equal = FALSE, conf.level = 0.95)
```
As the p-value < 0.05 the null hypothesis is rejected. (There is a difference in effects on tooth growth for dosages 1 and 2, dosage 2 having bigger effect)

--> Dosage matters: The more supplement, the more tooth growth